---
title: "Tarea 1 TS"
author: "Federico Daverio"
output:
  html_document:
    toc: yes
    toc_depth: 3
    number_sections: no
    theme: united
    highlight: tango
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
options(scipen=999) 
memory.limit(size=56000)
library(tidyverse)
library(stargazer)
library(Synth)
library(ggplot2)

setwd('C:/DAVE2/CIDE/4 semestre/Time series/Tarea_1')
```


# Exercise 1

## Video 1

No exercise

## Video 2

No exercise

## Video 3

### Square root of 729

```{r}
sqrt(729)
```
### Reminder of $109/25$

```{r}
109%%25
```
## Video 4
### Create a factor variable with 3 levels and 6 items
```{r}
fact<- factor(c("good", "bad", "ugly", "bad", "bad", "bad"))

class(fact)

nlevels(fact)

levels(fact)

```

## Video 5
### Assuming that you don't know the size of 'ab', get the first and the last element of the vector

```{r}
ab<- seq(1,293, by=7)

first <- ab[1]

last<- ab[length(ab)]

cat("First element: ", first)

cat(" Last element: ", last)

cat("or: ", ab[c(1, length(ab))])
```
## Video 6

### Craete a new directory in tyour computer. Open R-studio and set it as the working directory

```{r}
setwd('C:/DAVE2/CIDE/4 semestre/Time series/Tarea_1')
rm(list = ls())
```

Se borraron todas las variables así de tener en el ambiente solo *a* una vez creada

### Create a new variable 'a' and give it a value

```{r}
a<-324
```

### Save an image of your current session in "mySession.RData"

```{r}
save.image("mySession.RData")
```

### Delete all the objects in your current work space

```{r}
rm(list = ls())
```

### Load back "mySession.RData"

```{r}
load("mySession.RData")
```

## Video 7

### Create this pattern $1,1,1,2,2,3,4,4,4,4$

```{r}
c(rep(c(1,2,3), 3:1), rep(4,4))

```
Or:
```{r}
rep(1:4, times=c(3,2,1,4))
```


## Video 8

### Make a numeric vector that repeats the sequence of all odd numbers between 1 to 10 two times, with each number repeated thrice

```{r}
rep(seq(1,10, by=2),each=3, times=2)
```

## Video 10

Remove all the even numbers and the missing values from 'a'

```{r}
a<-c(1:5,1:4,NA)

a[!(a%%2==0)& !is.na(a)]

```

## Video 12
Create a vector(*v1*) using random numbers. Then convert that factor vector to its corresponding numeric vector (*v2*). All values in *v1* and *v2* should match.

```{r}
v1 <- factor(round(runif(30,30,50)))
v1
v2<- as.numeric(as.character(v1))
v2
v1==v2

```


## Video 13

### Create a list *l2* with a numeric c(1,2,3) and character vector c('a','b','c')

```{r}
l2<- list(c(1,2,3), c("a","b","c"))
```


### Give names to the vectors in list *l2*
```{r}
names(l2)<-c("numbers", "characters")
names(l2)
```
### Extract item "c" from the list using both notation ("$" and "[]")

```{r}
l2[[2]][3]
l2$characters[3]
```


## Video 14

### Get the non-repeating elements between a and b

```{r}
a<-c(11:15)
b<-c(13:17)

setdiff(union(a,b), intersect(a,b))

```

## Video 16

### Create a numeric vector of length 20, representing height of children entering a theme park. Let the heights lie between 1ft to 6 ft. If the height > 3ft, the entry is free else the entry is paid. 

```{r}
heights <- round(runif(20,1,6),1)

ifelse(heights>3, "free entry", "yuou need to pay")

```
## Video 17
### Create a for loop that pints alphabets from a to z

```{r}
for (i in letters){
  print(i)
}
```

## Video 18.1

No exercise

## Video 18.2

No exercise

## Video 19

### Exercise 1: Craete a dataframe with 5 columns and 100 rows. Each of the columns should contain random numbers between 1 and 10

```{r}
a<- round(runif(100, 1,10))
b<- round(runif(100, 1,10))
c<- round(runif(100, 1,10))
d<- round(runif(100, 1,10))
e<- round(runif(100, 1,10))

db <- data.frame(a,b,c,d,e)

head(db,10)
```
### Exercise 2: Insert 5 missing values in each of the 5 columns at random locations

```{r}
for (i in 1:5){
  db[sample(1:100,5),i]=NA
}

summary(db)

```
### Exercise 3: Get the row numbers that contain missing values in the first column.

```{r}
which(is.na(db$a))

```
### Exercise 4: Get the row number that contain missing values in any of the columns

```{r}
vec<- vector()
for (i in 1:ncol(db)){
  vec<-c(vec, which(is.na(db[,i])))
}

sort(unique(vec))
length(unique(vec))
```
### Exercise 5: Get the elements of 3rd and 5th column in the last 10 rows of db

```{r}
db[ 91:100,c(3,5)]
```

### Exercise 6: Sort the db such that the first column is sorted in ascending while second column is sorted in descending order.

```{r}
db[order(db$a,-db$b),]
```


## Video 20

### Challenge 1: Exctract the 25% of states (in rownames) with least arrests due to murder

```{r}
library(datasets)
USArrests

USArrests[USArrests$Murder<summary(USArrests$Murder)[2], 0]



```

Podemos tambien hacerlo de esta manera:

```{r}
USArrests[order(USArrests$Murder),]

q1<-nrow(USArrests)/4

USArrests[c(1:round(q1)),]
```

Notamos que hay una diferencia de un estado dado por la función *round*.

### Chaollenge 2: Extract the states whose Assalut is >200 or rape > 20

```{r}
USArrests[USArrests$Assault>200 | USArrests$Rape>20,0]
```

# Exercise 2.1

## Number 1
For poker_vector: On Monday you won 140 dls, Tuesday you lost 50 dls, Wednesday you won 20 dls, Thursday you lost 120 dls and Friday you won 240 dls.

```{r}
poker_vector<-c(140,-50,20,-120,240)
```


## Number 2
For roulette_vector: On Monday you lost 24 dls, Tuesday you lost 50 dls, Wednesday you won 100 dls, Thursday you lost 350 dls and Friday you won 10 dls


```{r}
roulette_vector<-c(-24,-50,100,-350,10)
```

## Number 3
Assign days as names of poker_vector

```{r}
names(poker_vector)<-c("Monday", "Tuesday", "Wednesday","Thursday", "Friday")

poker_vector
```

## Number 4
Assign days as names of roulette_vectors

```{r}
names(roulette_vector)<-c("Monday", "Tuesday", "Wednesday","Thursday","Friday")
roulette_vector
```
## Number 5
Assign the days of the week vector to a variable

```{r}
days_vector<-names(poker_vector)

days_vector
```
## Number 6
Calculate your total winnings (Hint - you have to do arithmetic calculations on vectors)

### Point A
How much has been your overall profit or loss per day of the week?

```{r}
for (i in days_vector){
  cat("Winning in ", i, (poker_vector[names(poker_vector)==i]+roulette_vector[names(roulette_vector)==i]), "\n")
}
```
O más sencillo:

```{r}
poker_vector+roulette_vector
```



### Point B
Have you lost money over the week in total?

```{r}
cat("My money at the end of the week: ", sum(poker_vector,roulette_vector))
```
Estamos por lo tanto perdiendo 84 dolares en la semana

O utilizando el hint:

```{r}
tot_vector<-poker_vector+roulette_vector
sum(tot_vector)
```

### Point C
Are you winning/losing money on poker or on roulette?

```{r}
cat("Winning/Losing on poker: ", sum(poker_vector), "\n")

cat("Winning/Losing on roulette poker: ", sum(roulette_vector))
```
Estamos ganando dinero en el poker y perdiendo dinero en la roulette

### Point D
To do it: add the poker_vector and roulette_vector and assign it to total_vector

```{r}
total_vector<-poker_vector + roulette_vector
sum(total_vector)
```

## Number 7
You wonder if there may be a very tiny chance you have lost money over the week in total? A function that helps you to answer this question is sum(). It calculates the sum of all elements of a vector.

### Point A
Calculate the total amount of money that you have won/lost with roulette and
poker and assign to the variable... total_roulette and total_poker

```{r}
total_poker<-sum(poker_vector)
total_poker
total_roulette<-sum(roulette_vector)
total_roulette
```

### Point B
Now that you have the totals for roulette and poker, you can easily calculate total_week (which is the sum of all gains and losses of the week).

```{r}
total_week<-sum(total_poker,total_roulette)
```

## Number 8
Check if your total gains in poker are higher than for roulette by using a comparison. Simply print out the result of this comparison. What do you conclude, should you focus on roulette or on poker?

```{r}
total_poker>total_roulette
```
Estamos ganando más con el poker. Si nuestra función de utilidad depende solo de las ganancias deberíamos concentrarnos en el poker.

## Number 9
Select the first three elements in poker_vector by using their names: "Monday", "Tuesday" and "Wednesday". Assign the result of the selection to poker_start.

```{r}
poker_start<- poker_vector[c("Monday", "Tuesday", "Thursday")]
poker_start
```
## Number 10
Calculate the average of the values in poker_start with the mean() function

```{r}
mean(poker_start)
```

# Exercis 2.2
Construct a data frame that describes the main characteristics of eight planets in our solar system.

* Name "Mercury", "Venus", "Earth", "Mars", "Jupiter", "Saturn", "Uranus", "Neptune"
```{r}
Name<-c("Mercury", "Venus", "Earth", "Mars", "Jupiter", "Saturn", "Uranus", "Neptune")
```

* The type of planet… "Terrestrial planet", "Terrestrial planet", "Terrestrial planet", "Terrestrial planet", "Gas giant", "Gas giant", "Gas giant", "Gas giant"

```{r}
Type<- c("Terrestrial planet", "Terrestrial planet", "Terrestrial planet", "Terrestrial planet", "Gas giant", "Gas giant", "Gas giant", "Gas giant")
```

* The planet's diameter relative to the diameter of the Earth. 0.382, 0.949, 1, 0.532, 11.209, 9.449, 4.007, 3.883

```{r}
Diameter <- c(0.382, 0.949, 1, 0.532, 11.209, 9.449, 4.007, 3.883)
```

* The planet's rotation across the sun relative to that of the Earth. 58.64, -243.02, 1, 1.03, 0.41, 0.43, -0.72, 0.67

```{r}
Rotation <- c(58.64, -243.02, 1, 1.03, 0.41, 0.43, -0.72, 0.67)
```

* if the planet has rings or not (TRUE or FALSE).. FALSE, FALSE, FALSE, FALSE, TRUE,TRUE, TRUE, TRUE)

```{r}
Rings <- c(FALSE, FALSE, FALSE, FALSE, TRUE,TRUE, TRUE, TRUE)
```

* Use the function data.frame() to construct a data frame. Pass the vectors name, type, diameter, rotation and rings as arguments to data.frame()

```{r}
Planets<- data.frame(Name, Type, Diameter,Rotation, Rings )
Planets

```

# Exercise 2.3.1

* Set a speed of 64

```{r}
speed<-64
```

* Code a while loop with the following characteristics:
 - The condition of the while loop should check if speed is higher than 30.
 - Inside the body of the while loop, print out "Slow down!".
 - Inside the body of the while loop, decrease the speed by 7 units and assign this new value to speed again.
 
```{r}
while(speed>30){
  print("Slow down!")
  speed= speed-7
}

print("Now your speed is legal!")
```
# Exercise 2.3.2 

* Add code to the for loop that loops over the elements of the linkedin vector:
* linkedin <- c(16, 9, 13, 5, 2, 17, 14)

```{r}
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
```

* If the vector element's value exceeds 10, print out "You're popular!".
* If the vector element's value does not exceed 10, print out "Be more visible!

```{r}
for(i in linkedin){
  ifelse(i>10, print("You're popular!"),print("Be more visible!"))
}


```
 
 